---
title: "Statistical assignment 4"
author: ""
date: "[add date here]"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = FALSE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r echo = TRUE}
library(tidyverse)
Data8 <- read_tsv("/Users/bessudnov/Documents/datan3_2020/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")

Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)

Stable <- read_tsv("/Users/bessudnov/Documents/datan3_2020/data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")

Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)

Data <- Data8 %>% left_join(Stable, "pidp")

rm(Data8, Stable)

Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )

    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points).

    ```{r}
Data %>%
        ggplot(aes(x = h_payn_dv)) +
        geom_freqpoly(bindwidth = 1) +
        xlab("Net monthly pay") +
        ylab("Number of respondents")
         
    ```

2. Line chart (20 points). The lines show the non-parametric association between age and monthly earnings for men and women.

    ```{r}
Data %>% 
            filter(!is.na(sex_dv)) %>%
            ggplot(aes(x = h_age_dv, y = h_payn_dv, linetype = sex_dv)) +
            geom_smooth(colour = "black") +
            xlim(16, 65) +
            xlab("Age") +
            ylab("Monthly earnings") +
            labs(linetype='Sex') 
    ```

3. Faceted bar chart (20 points).

    ```{r}
Data %>%
        filter(!is.na(placeBorn)) %>%
        filter(!is.na(sex_dv)) %>%
        group_by(placeBorn, sex_dv) %>%
        summarise(
                medianIncome = median(h_payn_dv, na.rm = TRUE)
        ) %>%
        ggplot(aes(y = medianIncome, x = sex_dv)) +
        geom_bar(stat = "identity", position = "dodge") +
        facet_wrap(~ placeBorn) +
        xlab("Sex") +
        ylab("Median monthly net pay")

    ```

4. Heat map (20 points).

    ```{r}
Data %>%
        filter(!is.na(placeBorn)) %>%
        filter(!is.na(h_gor_dv)) %>%
        group_by(h_gor_dv, placeBorn) %>%
            summarise(
                    meanAge = mean(h_age_dv, na.rm = TRUE)
            ) %>%
        ggplot(aes(x = h_gor_dv, y = placeBorn, fill = meanAge)) +
        geom_tile() +
        xlab("Region") +
        ylab("Country of birth") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        theme(panel.background = element_rect(fill = 'white')) +
        labs(fill = "Mean age")
    
    ```

5. Population pyramid (20 points).

    ```{r}

Data %>%
            filter(!is.na(sex_dv)) %>%
            filter(!is.na(h_age_dv)) %>%
            count(sex_dv, h_age_dv) %>%
            ggplot(aes(x = h_age_dv, y = ifelse(sex_dv == "female", n, -n), fill = sex_dv)) +
            geom_col() +
            coord_flip() +
            scale_fill_brewer(palette = "Set1") + 
            theme_bw() +
            xlab("Age") +
            ylab("n") +
            labs(fill = "Sex")

    ```